---
title: "Apache OpenNLP"
output: statsr:::statswithr_lab
author: Jacob John
---

#Digital Assignment 1

<div id="instructions">
CSE4022: Natural Language Processing
</div>
        
        
##Introduction
        
An interface to the Apache OpenNLP tools (version 1.5.3). The Apache OpenNLP library is a machine learning based toolkit for the processing of natural language text written in Java. It supports the most common NLP tasks, such as tokenization, sentence
segmentation, part-of-speech tagging, named entity extraction, chunking, parsing, and coreference resolution. See [OpenNLP](http://opennlp.apache.org/>) for more information


##Maxent_Chunk_Annotator
>Apache OpenNLP based chunk annotators

Generate an annotator which computes chunk annotations using the Apache OpenNLP Maxent chunker.

```{r}
require(rJava)
require(NLP)
require(openNLP)

## Requires package 'openNLPmodels.en' from the repository at
## <http://datacube.wu.ac.at>.
## Some text.
s <- paste(c("Pierre Vinken, 61 years old, will join the board as a ",
             "nonexecutive director Nov. 29.\n",
             "Mr. Vinken is chairman of Elsevier N.V., ","the Dutch publishing group."),
           collapse = "")
s <- as.String(s)
## Chunking needs word token annotations with POS tags.
sent_token_annotator <- Maxent_Sent_Token_Annotator()
word_token_annotator <- Maxent_Word_Token_Annotator()
pos_tag_annotator <- Maxent_POS_Tag_Annotator()
a3 <- annotate(s,
               list(sent_token_annotator,
                    word_token_annotator,
                    pos_tag_annotator))
annotate(s, Maxent_Chunk_Annotator(), a3)
annotate(s, Maxent_Chunk_Annotator(probs = TRUE), a3)
```


##Maxent_Entity_Annotator
>Apache OpenNLP based entity annotators

Generate an annotator which computes entity annotations using the Apache OpenNLP Maxent name finder.

```{r}
## Requires package 'openNLPmodels.en' from the repository at
## <http://datacube.wu.ac.at>.
require("NLP")
## Some text.
s <- paste(c("Pierre Vinken, 61 years old, will join the board as a ",
"nonexecutive director Nov. 29.\n",
"Mr. Vinken is chairman of Elsevier N.V., ",
"the Dutch publishing group."),
collapse = "")
s <- as.String(s)
## Need sentence and word token annotations.
sent_token_annotator <- Maxent_Sent_Token_Annotator()
word_token_annotator <- Maxent_Word_Token_Annotator()
a2 <- annotate(s, list(sent_token_annotator, word_token_annotator))
## Entity recognition for persons.
entity_annotator <- Maxent_Entity_Annotator()
entity_annotator
annotate(s, entity_annotator, a2)
## Directly:
entity_annotator(s, a2)
## And slice ...
s[entity_annotator(s, a2)]
## Variant with sentence probabilities as features.
annotate(s, Maxent_Entity_Annotator(probs = TRUE), a2)
```


##Maxent_POS_Tag_Annotator
>Apache OpenNLP based POS tag annotators

Generate an annotator which computes POS tag annotations using the Apache OpenNLP Maxent Part of Speech tagger.

```{r}
require("NLP")
## Some text.
s <- paste(c("Pierre Vinken, 61 years old, will join the board as a ",
"nonexecutive director Nov. 29.\n",
"Mr. Vinken is chairman of Elsevier N.V., ",
"the Dutch publishing group."),
collapse = "")
s <- as.String(s)
## Need sentence and word token annotations.
sent_token_annotator <- Maxent_Sent_Token_Annotator()
word_token_annotator <- Maxent_Word_Token_Annotator()
a2 <- annotate(s, list(sent_token_annotator, word_token_annotator))
pos_tag_annotator <- Maxent_POS_Tag_Annotator()
pos_tag_annotator
a3 <- annotate(s, pos_tag_annotator, a2)
a3
## Variant with POS tag probabilities as (additional) features.
head(annotate(s, Maxent_POS_Tag_Annotator(probs = TRUE), a2))
## Determine the distribution of POS tags for word tokens.
a3w <- subset(a3, type == "word")
tags <- sapply(a3w$features, `[[`, "POS")
tags
table(tags)
## Extract token/POS pairs (all of them): easy.
sprintf("%s/%s", s[a3w], tags)
## Extract pairs of word tokens and POS tags for second sentence:
a3ws2 <- annotations_in_spans(subset(a3, type == "word"),
subset(a3, type == "sentence")[2L])[[1L]]
sprintf("%s/%s", s[a3ws2], sapply(a3ws2$features, `[[`, "POS"))
```


##Maxent_Sent_Token_Annotator
>Apache OpenNLP based sentence token annotators

Generate an annotator which computes sentence annotations using the Apache OpenNLP Maxent sentence detector.

```{r}
require("NLP")
## Some text.
s <- paste(c("Pierre Vinken, 61 years old, will join the board as a ",
"nonexecutive director Nov. 29.\n",
"Mr. Vinken is chairman of Elsevier N.V., ",
"the Dutch publishing group."),
collapse = "")
s <- as.String(s)
sent_token_annotator <- Maxent_Sent_Token_Annotator()
sent_token_annotator
a1 <- annotate(s, sent_token_annotator)
a1
## Extract sentences.
s[a1]
## Variant with sentence probabilities as features.
annotate(s, Maxent_Sent_Token_Annotator(probs = TRUE))
```

##Maxent_Word_Token_Annotator
>Apache OpenNLP based word token annotators

Generate an annotator which computes word token annotations using the Apache OpenNLP Maxent tokenizer

```{r}
require("NLP")
## Some text.
s <- paste(c("Pierre Vinken, 61 years old, will join the board as a ",
"nonexecutive director Nov. 29.\n",
"Mr. Vinken is chairman of Elsevier N.V., ",
"the Dutch publishing group."),
collapse = "")
s <- as.String(s)
## Need sentence token annotations.
sent_token_annotator <- Maxent_Sent_Token_Annotator()
a1 <- annotate(s, sent_token_annotator)
word_token_annotator <- Maxent_Word_Token_Annotator()
word_token_annotator
a2 <- annotate(s, word_token_annotator, a1)
a2
## Variant with word token probabilities as features.
head(annotate(s, Maxent_Word_Token_Annotator(probs = TRUE), a1))
## Can also perform sentence and word token annotations in a pipeline:
a <- annotate(s, list(sent_token_annotator, word_token_annotator))
head(a)
```


##Parse_Annotator
>Apache OpenNLP based parse annotator

Generate an annotator which computes Penn Treebank parse annotations using the Apache OpenNLP chunking parser for English.

```{r}
## Requires package 'openNLPmodels.en' from the repository at
## <http://datacube.wu.ac.at>.
require("NLP")
## Some text.
s <- paste(c("Pierre Vinken, 61 years old, will join the board as a ",
"nonexecutive director Nov. 29.\n",
"Mr. Vinken is chairman of Elsevier N.V., ",
"the Dutch publishing group."),
collapse = "")
s <- as.String(s)
## Need sentence and word token annotations.
sent_token_annotator <- Maxent_Sent_Token_Annotator()
word_token_annotator <- Maxent_Word_Token_Annotator()
a2 <- annotate(s, list(sent_token_annotator, word_token_annotator))
parse_annotator <- Parse_Annotator()
## Compute the parse annotations only.
p <- parse_annotator(s, a2)
## Extract the formatted parse trees.
ptexts <- sapply(p$features, `[[`, "parse")
ptexts
## Read into NLP Tree objects.
ptrees <- lapply(ptexts, Tree_parse)
ptrees
```